hbm_paper_results

### 2024-03-13 paper_replication_results.Rmd
cleaned upgit add paper*

@@

```{r}
cent.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
ggplot(aes(x=perm, fill=bigger_group)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~pheno_cat)
```

@@

ct.brain.sum.plt <- ct.t.tests_to_plt.sum %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
            aes(fill=percent_cf.gamlss_better))  +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications where ComBatLS is Better")
```

@@

Plot raw values from random permutation

```{r, eval=FALSE}
raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE) %>% unlist()

diff.df.rand <- fread(raw_files[3])

diff.df.rand <- diff.df.rand %>%
  mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|perm|-", "", Source_File))

global_abs.diffs <- diff.df.rand %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", pheno_list)) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
    mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         abs.diff = abs.diff*100,
         pheno_cat = factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat

dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
all.abs.err.plt <- global_abs.diffs %>%
  dplyr::filter(dataset != "cf") %>%
  ggplot() +
  geom_boxplot(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=abs.diff, fill=dataset), outlier.size=.1) +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Absolulte Centile Error", x = "Feature") +
  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
  facet_wrap(~dataset, nrow=3, labeller = as_labeller(dataset_names), scales="free_y") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="none")
ggsave("figs/all_abs_err_plt_wlabels.jpeg", all.abs.err.plt, width=16, height=9, units="in")
```

@@

```{r}
cent.subj.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("lightblue", "red"))
```

@@

t.test.cent.sex.test.med <- cent.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ "")) %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")

@@

sex_bias_density_plt <- cent.sex.test.med %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
              data = t.test.cent.sex.test.med) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")

cent.sex.test.med %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss" | dataset == "cf.lm") %>%
  ggplot() +
  geom_boxplot(aes(y=median_of_median_sex_diff*100, x=dataset, color=dataset)) +
  geom_hline(yintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat, scales="free") +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos*100, y=10.2),
              data = t.test.cent.sex.test.med) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Centile Errors Within Brain Feature Types", y="Density Across 100 Replications")
```

@@

all_cent_sum.df %>%
  ggplot(aes(x=dataset, y=female_ratio)) +
  geom_boxplot(aes(color=dataset)) +
  # geom_dotplot(binaxis = "y", stackdir="center") +
  facet_wrap(~source, scales="free_y") +
  #geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
  # geom_text(aes(label = sig_fdr, y = max_val + 0.01),
  #             data = all_t_test.df) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
  theme_linedraw()
```

@@

by site
```{r}
perm_subj.df2.by.site <- perm_subj.df2 %>%
  group_by(perm, dataset, sim.site) %>%
  summarise(across(contains("mean"), list(site = mean)))

perm_subj.df2.by.site %>%
    mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(x=sim.site, y=mean_cent_abs.diff_site, fill=dataset)) +
  facet_wrap(~ dataset, scales="free")

perm_subj.df2.by.site %>%
  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset)) +
  facet_wrap(~ dataset, scales="free")

perm_subj.df2.by.site %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(x=sim.site, y=mean_cent_diff_site, fill=dataset))

perm_subj.df2.by.site %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(x=sim.site, y=mean_centile_site, fill=dataset))

perm_subj.df2.by.site %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  pivot_wider(id_cols=c(perm, sim.site), names_from=dataset, values_from=mean_centile_site, names_prefix="site_mean_cent_") %>%
  mutate(delta_mean = site_mean_cent_cf.gam - site_mean_cent_cf.gamlss) %>%
  mutate(sim.site = factor(sim.site, levels=c("Site_B", "Site_A", "Site_C"), ordered=TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(x=sim.site, y=delta_mean))
```

@@

```{r}
noext.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
noext.subj.sex.test.df <- bind_rows(noext.subj.sex.test.list)
```

```{r}
noext.subj.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("lightblue", "red"))
```

@@

```{r}
z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=perm, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~comp)
```

@@

```{r}
z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
ggplot(aes(x=perm, fill=bigger_group)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~pheno_cat)
```

@@

```{r}
z.diffs.v.plot <- perm_subj.df %>%
  ggplot() +
  geom_violin(aes(x=dataset, y=mean_z_abs.diff, fill=perm), scale="count") +
  labs(y= "Mean Absolute Z-Score Error", x = "ComBat Configurations") +
  scale_x_discrete(labels=c(dataset_names)) +
  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
  ggtitle("Within-subject mean magnitude of error in z scores across ComBat \nconfigurations replicated 10x")
z.diffs.v.plot
# ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```

@@

```{r}
z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
z.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
```

@@

```{r}
z.subj.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_fill_manual(values = c("lightblue", "red"))
```

>>>>

**need to prettify table for supplement**

@@

```{r}
#look at combatGAM vs combatLS
cent.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
ggplot(aes(x=perm, fill=bigger_group)) +
  geom_bar() +
  # geom_text(stat='count', aes(label=..count..)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~pheno_cat)

cf.gam_worse <- cent.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam" & bigger_group == "cf.gamlss") %>%
  group_by(perm) %>%
  summarise(count = n())

mean(cf.gam_worse$count)
IQR(cf.gam_worse$count)
```

@@

ct.brain.sum.plt <- ct.t.tests_to_plt.sum %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
            aes(fill=percent_cf.gamlss_better))  +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  scale_fill_gradient(low="#605210", high= "#e6d16d",  name = "% Replications") +
  labs(title="Replications where ComBatLS recapitulates cortical thickness centiles \nsignificantly better than ComBat-GAM")
```

@@

@@

```{r}
cent.subj.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  #geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values = c("lightblue", "red"))
```

@@

t.test.cent.sex.test.med <- cent.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""))

t.test.cent.sex.test.med.filt <- t.test.cent.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")

@@

sex_bias_density_plt <- cent.sex.test.med %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med.filt$dataset == "cf.gam", 10.2, 5.5)),
              data = t.test.cent.sex.test.med.filt) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")

cent.sex.test.med %>%
  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
              data = t.test.cent.sex.test.med) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
```

@@

@@

@@

```{r}
noext.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
noext.sex.test.df <- bind_rows(noext.sex.test.list)
```

simple bar-plot to count regions w sigificant sex-differences

```{r}
noext.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  #geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values = c("lightblue", "red"))
```

Distribution of error medians w/in each pheno category
```{r}
noext.sex.testmed <- noext.sex.test.df %>%
   mutate(pheno_cat = factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  group_by(dataset, perm, pheno_cat) %>%
  summarise(median_of_median_sex_diff = median(median_sex_diff))

next.t.test.cent.sex.test.med <- noext.sex.testmed %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(cent.sex.test.med$dataset))*length(unique(cent.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""))

next.t.test.cent.sex.test.med %>%
  ggplot() +
  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")

noext.sex.testmed %>%
  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff*100, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos*100, y=ifelse(t.test.cent.sex.test.med$dataset == "cf.gam", 10.2, 5.5)),
              data = next.t.test.cent.sex.test.med) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Centile Errors", y="Density Across 100 Replications")
```

@@

```{r}
z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=perm, fill=bigger_group)) +
  theme(axis.text.x = element_blank()) +
  facet_grid(pheno_cat~comp, scales="free_y", labeller = labeller(comp = as_labeller(comp_names)))
```

@@

```{r}
z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
ggplot(aes(x=perm, fill=bigger_group)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~pheno_cat)

z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE & comp =="cf.gamlss_cf.gam") %>%
ggplot(aes(x=perm, fill=bigger_group)) +
  geom_bar(position='fill') +
  geom_hline(yintercept = .5) +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~pheno_cat)
```

@@

```{r}
z.diffs.v.plot <- perm_subj.df %>%
  ggplot() +
  geom_violin(aes(x=dataset, y=mean_z_abs.diff, color=perm), scale="count") +
  labs(y= "Mean Absolute Z-Score Error", x = "ComBat Configurations") +
  scale_x_discrete(labels=c(dataset_names)) +
  theme(legend.position = "none") +
  ggtitle("Within-subject mean magnitude of error in z scores across ComBat \nconfigurations replicated 10x")
z.diffs.v.plot
# ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```

@@

```{r}
z.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
z.sex.test.df <- bind_rows(z.sex.test.list)
```

@@

```{r}
z.sex.test.df %>%
  mutate(sig_fdr = factor(sig_fdr, ordered=TRUE, levels=c("TRUE", "FALSE"))) %>%
  ggplot(aes(x=perm, fill=sig_fdr)) +
  geom_bar() +
  #geom_text(stat='count', aes(label=..count..)) +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values = c("lightblue", "red"))
```

Distribution of error medians w/in each pheno category
```{r}
z.sex.test.med <- z.sex.test.df %>%
  mutate(pheno = gsub(".z", "", pheno)) %>%
   mutate(pheno_cat = factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  group_by(dataset, perm, pheno_cat) %>%
  summarise(median_of_median_sex_diff = median(median_sex_diff))

t.test.z.sex.test.med <- z.sex.test.med %>%
  group_by(dataset, pheno_cat) %>%
  summarise(t.test.p = tidy(t.test(median_of_median_sex_diff))$p.value,
            t.test.est = tidy(t.test(median_of_median_sex_diff))$estimate,
            x_pos = median(median_of_median_sex_diff)) %>% #test median of each distribution against 0
  mutate(p.val_fdr = p.adjust(t.test.p, method="fdr", n = length(unique(z.sex.test.med$dataset))*length(unique(z.sex.test.med$pheno_cat))),
                  sig_fdr = case_when(p.val_fdr < 0.001 ~ "***",
                               p.val_fdr >= 0.001 & p.val_fdr < 0.01 ~ "**",
                               p.val_fdr >= 0.01 & p.val_fdr < 0.05 ~ "*",
                               p.val_fdr >= 0.05 ~ ""))

t.test.z.sex.test.med.filt <- t.test.z.sex.test.med %>% dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss")

t.test.z.sex.test.med %>%
  ggplot() +
  geom_col(aes(x=pheno_cat, y=sig_fdr, fill=dataset), position="dodge")

z.sex_bias_density_plt <- z.sex.test.med %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med.filt$dataset == "cf.gam", 350, 120)),
              data = t.test.z.sex.test.med.filt) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")

z.sex.test.med %>%
  #dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  ggplot() +
  geom_density(aes(x=median_of_median_sex_diff, fill=dataset, color=dataset), alpha=0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed", color="#3F3F3F") +
  facet_wrap(~pheno_cat) +
  theme_linedraw() +
  geom_text(aes(label = sig_fdr, x= x_pos, y=ifelse(t.test.z.sex.test.med$dataset == "cf.gam", 350, 120)),
              data = t.test.z.sex.test.med) +
  scale_fill_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  theme(legend.position = "bottom")+
  labs(x="Medians of Sex-Differences in Z Errors", y="Density Across 100 Replications")
```

plot subj. with averages in extremes 
```{r}
#recalc subject's true mean zs
raw_mean_z <- perm_subj.df %>%
  dplyr::filter(dataset=="cf") %>%
  mutate(mean_z = mean_z - mean_z_diff,
        VolGlob_mean_z = VolGlob_mean_z - VolGlob_mean_z_diff,
        VolReg_mean_z = VolReg_mean_z - VolReg_mean_z_diff,
        CT_mean_z = CT_mean_z - CT_mean_z_diff,
        SA_mean_z = SA_mean_z - SA_mean_z_diff) %>%
  dplyr::select(!c(contains("diff"), contains("centile"))) %>%
  mutate(dataset="raw")

raw_mean_z %>%
  dplyr::filter(perm=="perm-001") %>%
  dplyr::filter(SA_mean_z > 1.5) %>% 
  nrow()

perm_subj.df.z <- full_join(perm_subj.df, raw_mean_z)

perm_subj.df.z._long <- perm_subj.df.z %>%
  rename(allpheno_mean_z = mean_z) %>%
  pivot_longer(cols=contains("mean_z"), names_to="pheno_cat", values_to="mean_z") %>%
  mutate(pheno_cat = gsub("_mean_z", "", pheno_cat)) %>%
  mutate(pheno_cat = case_when(pheno_cat == "VolGlob" ~ "Global Volume",
                               pheno_cat == "VolReg" ~ "Regional Volume",
                               pheno_cat == "CT" ~ "Regional CT",
                               pheno_cat == "SA" ~ "Regional SA"))

high_z.sum <- perm_subj.df.z._long %>%
  dplyr::filter(pheno_cat != "allpheno") %>%
  dplyr::filter(mean_z > 1.5) %>%
  group_by(dataset, perm, pheno_cat) %>%
  summarise(
    male_ratio = sum(sex == "Male") / n(),
    female_ratio = sum(sex == "Female") / n()
  )

low_z.sum <- perm_subj.df.z._long %>%
  dplyr::filter(pheno_cat != "allpheno") %>%
  dplyr::filter(mean_z < -1.5) %>%
  group_by(dataset, perm, pheno_cat) %>%
  summarise(
    male_ratio = sum(sex == "Male") / n(),
    female_ratio = sum(sex == "Female") / n()
  )

ext_z.sum <- perm_subj.df.z._long %>%
  dplyr::filter(pheno_cat != "allpheno") %>%
  dplyr::filter(mean_z < -1.5 | mean_z > 1.5) %>%
  group_by(dataset, perm, pheno_cat) %>%
  summarise(
    male_ratio = sum(sex == "Male") / n(),
    female_ratio = sum(sex == "Female") / n()
  )
```

```{r}
#get raw vals for horizontal lines
high_z.sum_raw <- high_z.sum %>%
  dplyr::filter(dataset == "raw" & perm=="perm-001")

low_z.sum_raw <- low_z.sum %>%
  dplyr::filter(dataset == "raw" & perm=="perm-001")

ext_z.sum_raw <- ext_z.sum %>%
  dplyr::filter(dataset == "raw" & perm=="perm-001")

z.high_t_test.df <- lapply(split(high_z.sum, high_z.sum$pheno_cat), get_t_tests, raw_df=high_z.sum_raw) %>% bind_rows()
z.low_t_test.df <- lapply(split(low_z.sum, low_z.sum$pheno_cat), get_t_tests, raw_df=low_z.sum_raw) %>% bind_rows()
z.ext_t_test.df <- lapply(split(ext_z.sum, ext_z.sum$pheno_cat), get_t_tests, raw_df=ext_z.sum_raw) %>% bind_rows()

#merge dfs
all_z_sum.df <- gdata::combine(high_z.sum, low_z.sum, ext_z.sum) %>%
  mutate(source = gsub("_z.sum", "", source))
all_z.sum_raw <- all_z_sum.df %>%
  dplyr::filter(dataset == "raw" & perm=="perm-001")

z.all_t_test.df <- gdata::combine(z.high_t_test.df, z.low_t_test.df, z.ext_t_test.df) %>%
  mutate(source = gsub("_t_test.df", "", source)) %>%
  mutate(source = gsub("z.", "", source))

#scale by raw value
all_z_sum.df_scaled <- all_z_sum.df %>%
  mutate(dataset = factor(dataset, levels=c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered=TRUE)) %>%
  arrange(dataset) %>%
  group_by(source, pheno_cat) %>%
  mutate(female_ratio_sc = female_ratio/first(female_ratio),
         female_ratio_diff = female_ratio - first(female_ratio))

#plots!
all_z_sum.df %>%
  dplyr::filter(dataset != "raw") %>%
  ggplot(aes(x=dataset, y=female_ratio)) +
  geom_boxplot(aes(color=dataset)) +
  # geom_dotplot(binaxis = "y", stackdir="zer") +
  facet_grid(pheno_cat~source, scales="free_y") +
  geom_hline(data=all_z.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
              data = z.all_t_test.df) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"),  name = "ComBat Configuration", labels=dataset_names) + 
  theme_linedraw()

z_ext_list <- c(ext = ">1.5 and <-1.5", high = ">1.5", low = "<-1.5")
